---
title: "Homework 4"
author: "Ahmed_Drammeh"
date: "4/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Obesity Hospitalization Data in England

This dataset contains informaion regarding **finished admissions episodes** (FAE) for people in the United Kingdom who were admitted due to obesity issues. The dataset in particular is part of a larger grouped dataset on the same excel sheet, and requires tedious navigation inorder to separate the **FAE** information from the other redundant tables. The **FAE** episodes involving obese patients are categorized into age groups ranging from:

* Under 16
* 16 - 24
* 25 - 34
* 35 - 44
* 45 - 54
* 55 - 64
* 65 - 74
* 75 & Over

Furthermore, these hospitalizations were recorded from 2002-2003 fiscal year all the way through 2012-2013. On the RAW excel spreadsheet, the dataset can be identified as *Table 7.2: Finished Admission Episodes with a primary diagnosis of obesity, by age group, 2002/03 to 2012/13*
Inorder for any data analysis to be performed on the obesity information, the data had to be cleaned first. The data was imported into R-studio by skipping the first 37 rows of the spreadsheet, since these rows include data from **Table 7.1**, and not **7.2**. Apart from skipping the first 37 rows, the number of rows containing the data of interest had to be inputed within the read.csv file, and is inlcuded within the script. There are 13 rows and 12 columns within the dataset.
The next paragraph describes the data cleaning and tidying effort:


## Data cleaning & tidying

The imported medical obesity dataset was poorly organized and the column and row names needed to be tidied and renamed. There are two ways to approach this. The first is with conventional R methodolgies of assigning new column names to an object, for example in our case( col.names), and then use this new object as an argument within the **names()** function to rename the column. The sequence of scripts below show this relationship. The second approach utilizes the **DPYLR** package, specifically the *rename()*, *select()*, and *slice()* functions to change the column names while also removing unnecessary empty rows. The second approach combined with the pipe operator makes for a much easier tidying method. The dataset can be further tidied, by removing organizing the different ages into one column called *Age groups*,making it both ore visually appealing to look at aswell as easier for statistical analysis. Under the tidyverse package, the **gather()** is a great function to help consodlidate data from wide to long format, and that has been utilized in this portion of the data tidying. Utilizing the gather function helped change the data format from wide to long, while also removing all the *NA* values which were present.

```{r Obesity1, echo=TRUE}
library(tidyverse)
library(dplyr)

Obesity1<- read.csv("obes-phys-acti-diet-eng-2014-tab_csv.csv",header =TRUE, stringsAsFactors= FALSE, skip=37, nrows= 13)
str(Obesity1)
colnames(Obesity1)
Obesity2<-Obesity1 %>% rename(Fiscal_Year= X, "Under 16" = Under.16,
                             "16 to 24"= X16.24, "25 to 34" = X25.34,
                             "35 to 44" = X35.44,"45 to 54" = X45.54,
                             "55 to 64" = X55.64,"65 to 74" = X65.74,
                             "75 & Over" = X75.and.over) %>%
  select(1:10) %>% slice(-1)
Obesity3<-Obesity2%>% 
  slice(-12) %>%
  gather(key="Age_Groups", value="Obesity_cases", -1,-2, na.rm=TRUE)

## Used global substitution function to change the character strings into numeric inroder to analyze them. dplyr as.numeric did not work well with numbers withi rows that have a comma(,) in them i.e.2,567

Obesity3$Total<-as.numeric(as.character(gsub(",", "",Obesity3$Total)))
Obesity3$Obesity_cases<-as.numeric(as.character(gsub(",", "",Obesity3$Obesity_cases)))

```

## Descriptive statistics

Three types of data visualization were explored. The first was a simple **scatter plot** to show the drastic increase in hospitalization cases from fiscal year **2002/2003** - **2012/2013**. From the scatter plot, an almost standardized distribution can be seen regarding the degree of obesity cases. The age groups **35-44 & 45-54** have the highest number of hospitalizations per year throughout the enitre study and is denoted by the higher peaks within the graph. However, using a slightly different graphical theme (changing the transparency of the markers) does a better job of highlighting the differences.

The second type of visulazation was a histogram, which didnt really do a good job of highlighting the differences between Age groups for the overall population at large.

The scripts below show some statistical analysis of the Obesity hospitalization data. The major findings include measuremenets of central tendancy (**mean**, **standard deviation** and **coefficient of variation**), aswell as some 5 number summary including the **min**, **max** and **median**. Looking at the overall data from a global perspective, it can be seen that the age groups **35-44 & 45-54** have the highest number of hospitalizations per year due to obesity( **Max = 3277 & 3581, and on average = 1723 & 1740**) over the time frames between *2002/2003* - *2012/2013*. This makes sense, since metabolic rates diminish significantly around those ages. Also people tend to be married and care less about what they eat or look like arounfd those ages.
The age groups with the lowest hospitalizations and obesity rates are **75 & Over( max = 115 and on average = 59)**.

```{r Obesity3, echo=TRUE}
library(tidyverse)
Obesity3 %>%
  ggplot(aes(y=Obesity_cases, x=Age_Groups, color=Fiscal_Year)) + geom_point(size=6) +coord_cartesian(ylim=c(40,4000))

## Changing the size and tranparency of the marker's help magnify the impact of the fiscal year on Obesity hospitalizations.

Obesity3 %>%
  ggplot(aes(y=Obesity_cases, x=Age_Groups, size=Fiscal_Year,color=Fiscal_Year, alpha=Fiscal_Year)) + geom_point() +coord_cartesian(ylim=c(40,4000)) + ggtitle("Hospitalizations due to Obesity")

## Histogram does not do a good job of easily discriminating between the Age groups.

Obesity3 %>%
  ggplot(aes(x =Obesity_cases, fill=Age_Groups )) + geom_histogram(binwidth = 500) + coord_cartesian(xlim=c(40,4000)) + ggtitle("Hospitalizations due to Obesity")

## Boxplots are another means to decribe the data in a way that helps easily visualize trends.

Obesity3 %>%
  ggplot(aes(y =Obesity_cases, x=Age_Groups, fill=Age_Groups )) + geom_boxplot() + coord_cartesian(ylim=c(40,4000)) + ggtitle("Hospitalizations due to Obesity")

Obesity4<-Obesity3 %>% 
  group_by(Age_Groups) %>% 
  summarise(Obesity_Min = min(Obesity_cases),
            Obesity_mean = round(mean(Obesity_cases)),
            Obesity_max = max(Obesity_cases),
            Std.dev = sd(Obesity_cases)) %>%
  mutate(COV = round((Std.dev/Obesity_mean),digits =2))
head(Obesity4)
## To keep to the original format of the data, i assigned a sorting mechanism to an object called *reorder*, and then used the slice(match()) functions to rearrange them in my desired order. R naturally arranges rows based on numerics first, and strings second.
reorder<-c("Under 16", "16 to 24", "25 to 34", "35 to 44", "45 to 54", "55 to 64", "65 to 74", "75 & Over")
Obesity4%>% slice(match(reorder, Age_Groups))

## Use of a histogram however, does a good job of showing the differences in Age groups when the data is summarized using measures of central tendancy.

## Average numer of obesity hospitalization when the fiscal year is lumped together.
Obesity4 %>%
  ggplot(aes(x =Obesity_mean, fill=Age_Groups, group=Age_Groups )) + geom_histogram(binwidth = 50) + coord_cartesian(xlim=c(40,2000)) + ggtitle("Average hospitalizations of Obesity cases")

## Standard deviation of the hospitalization rate.

Obesity4 %>%
  ggplot(aes(x =Std.dev, fill=Age_Groups, group=Age_Groups )) + geom_histogram(binwidth = 50) + coord_cartesian(xlim=c(40,1500)) + ggtitle("Standard devition of Obesity data")

## Coefficient of variation the hospitalization rate.

Obesity4 %>%
  ggplot(aes(x =COV, fill=Age_Groups, group=Age_Groups )) + geom_histogram(binwidth = .02) + coord_cartesian(xlim=c(0,1)) + ggtitle("Cofficient of Variation of Obesity data")

## Perhaps the best representation of the summaried data can be described in the form of a barplot or a boxplot

Obesity4 %>%
  ggplot(aes(x=Age_Groups, y=COV)) +geom_bar(stat="identity", color="black", fill="light blue") + ggtitle("Cofficient of Variation of Obesity data")

## Avergae number of hospitalizations

Obesity4 %>%
  ggplot(aes(x=Age_Groups, y=Obesity_mean)) +geom_bar(stat="identity", color="black", fill="light blue") +  ggtitle("Average hospitalizations of Obesity cases")

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
