---
title: "Homework 5"
author: "Ahmed_Drammeh"
date: "4/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Obesity Hospitalization Data in England

This dataset contains informaion regarding **finished admissions episodes** (FAE) for people in the United Kingdom who were admitted due to obesity issues. The dataset in particular is part of a larger grouped dataset on the same excel sheet, and requires tedious navigation inorder to separate the **FAE** information from the other redundant tables. The **FAE** episodes involving obese patients are categorized into age groups ranging from:

* Under 16
* 16 - 24
* 25 - 34
* 35 - 44
* 45 - 54
* 55 - 64
* 65 - 74
* 75 & Over

Furthermore, these hospitalizations were recorded from 2002-2003 fiscal year all the way through 2012-2013. On the RAW excel spreadsheet, the dataset can be identified as *Table 7.2: Finished Admission Episodes with a primary diagnosis of obesity, by age group, 2002/03 to 2012/13*
Inorder for any data analysis to be performed on the obesity information, the data had to be cleaned first. The data was imported into R-studio by skipping the first 37 rows of the spreadsheet, since these rows include data from **Table 7.1**, and not **7.2**. Apart from skipping the first 37 rows, the number of rows containing the data of interest had to be inputed within the read.csv file, and is inlcuded within the script. There are 13 rows and 12 columns within the dataset.
The next paragraph describes the data cleaning and tidying effort:


## Data cleaning & tidying

The imported medical obesity dataset was poorly organized and the column and row names needed to be tidied and renamed. There are two ways to approach this. The first is with conventional R methodolgies of assigning new column names to an object, for example in our case( col.names), and then use this new object as an argument within the **names()** function to rename the column. The sequence of scripts below show this relationship. The second approach utilizes the **DPYLR** package, specifically the *rename()*, *select()*, and *slice()* functions to change the column names while also removing unnecessary empty rows. The second approach combined with the pipe operator makes for a much easier tidying method. The dataset can be further tidied, by removing organizing the different ages into one column called *Age groups*,making it both ore visually appealing to look at aswell as easier for statistical analysis. Under the tidyverse package, the **gather()** is a great function to help consodlidate data from wide to long format, and that has been utilized in this portion of the data tidying. Utilizing the gather function helped change the data format from wide to long, while also removing all the *NA* values which were present.

```{r Obesity1, echo=TRUE}
library(tidyverse)
library(dplyr)

Obesity1<- read.csv("obes-phys-acti-diet-eng-2014-tab_csv.csv",header =TRUE, stringsAsFactors= FALSE, skip=37, nrows= 13)
str(Obesity1)
colnames(Obesity1)
Obesity2<-Obesity1 %>% rename(Fiscal_Year= X, "Under 16" = Under.16,
                             "16 to 24"= X16.24, "25 to 34" = X25.34,
                             "35 to 44" = X35.44,"45 to 54" = X45.54,
                             "55 to 64" = X55.64,"65 to 74" = X65.74,
                             "75 & Over" = X75.and.over) %>%
  select(1:10) %>% slice(-1)
Obesity3<-Obesity2%>% 
  slice(-12) %>%
  gather(key="Age_Groups", value="Obesity_cases", -1,-2, na.rm=TRUE)

## Used global substitution function to change the character strings into numeric inroder to analyze them. dplyr as.numeric did not work well with numbers withi rows that have a comma(,) in them i.e.2,567

Obesity3$Total<-as.numeric(as.character(gsub(",", "",Obesity3$Total)))
Obesity3$Obesity_cases<-as.numeric(as.character(gsub(",", "",Obesity3$Obesity_cases)))

```

## Bivariate Statistics

In homework 4, we looked at Univariate staistics and showed relationship between one variable and its influence on the cases of hospitalizations due to obesity. We saw that certain graphs did a better job of describing the data visually. For example using boxplot and barplots helped really capture how the mean cases of hospitalizations were in the age groups **35-44 & 45-54** 
We can also look at multiple varibles and see there relationship to each other. Utilizng facet wrap as a third variable in a bivariate study really brings out a different level of visual understanding. Looking at a few cases regarding using histograms and scatterplots, use of facet wrap as can be seen in the figures below isolate each set of independent variables in its own graph. It can be easily seen that the number of hospitalizations have steadily increased from fiscal year **2002/2003** to **2012/2013**. You can also see the intial progression of cases within age groups **Under 16 - 45-54**, and then a steady degression from **45-54 - 75 and Over**
Furthermore, using multiple geom_point() plots allowed the comparison of **Obesity cases** vs **Age groups**, while superimposing summary statistics data such as (Minimum, Mean, std.dev, COV and Max) as a variables.


```{r Obesity3, echo=TRUE}
library(tidyverse)
Obesity3 %>%
  ggplot(aes(y=Obesity_cases, x=Age_Groups, color=Fiscal_Year)) + geom_point(size=6) +coord_cartesian(ylim=c(40,4000))

## Changing the size and tranparency of the marker's help magnify the impact of the fiscal year on Obesity hospitalizations.

Obesity3 %>%
  ggplot(aes(y=Obesity_cases, x=Age_Groups, size=Fiscal_Year,color=Fiscal_Year, alpha=Fiscal_Year)) + geom_point() +coord_cartesian(ylim=c(40,4000)) + ggtitle("Hospitalizations due to Obesity")

## Combining Facet wrap with the scatter plot by using Age Groups as a third variable.

Obesity3 %>%
  ggplot(aes(y=Obesity_cases, x=Age_Groups, size=Fiscal_Year,color=Fiscal_Year, alpha=Fiscal_Year)) + geom_point() +coord_cartesian(ylim=c(40,4000)) + ggtitle("Hospitalizations due to Obesity") + facet_wrap(~Age_Groups, nrow=2) + theme(axis.text.x = element_text(angle=90))


## Histogram does not do a good job of easily discriminating between the Age groups.

Obesity3 %>%
  ggplot(aes(x =Obesity_cases, fill=Age_Groups )) + geom_histogram(binwidth = 500) + coord_cartesian(xlim=c(40,4000)) + ggtitle("Hospitalizations due to Obesity")

##Facet wrapping of the histogram helps deconvolute the data and establishes a pattern.

Obesity3 %>%
  ggplot(aes(x =Obesity_cases, fill=Age_Groups )) + geom_histogram(binwidth = 500) + coord_cartesian(xlim=c(40,4000)) + ggtitle("Hospitalizations due to Obesity") + facet_wrap(~Fiscal_Year,nrow=3)

## Boxplots are another means to decribe the data in a way that helps easily visualize trends.

Obesity3 %>%
  ggplot(aes(y =Obesity_cases, x=Age_Groups, fill=Age_Groups )) + geom_boxplot() + coord_cartesian(ylim=c(40,4000)) + ggtitle("Hospitalizations due to Obesity")

Obesity3 %>%
  ggplot(aes(y =Obesity_cases, x=Age_Groups, fill=Age_Groups )) + geom_boxplot() + coord_cartesian(ylim=c(40,4000)) + ggtitle("Hospitalizations due to Obesity") + facet_wrap(~Fiscal_Year, nrow=3)


Obesity4<-Obesity3 %>% 
  group_by(Age_Groups) %>% 
  summarise(Obesity_Min = min(Obesity_cases),
            Obesity_mean = round(mean(Obesity_cases)),
            Obesity_max = max(Obesity_cases)) %>%
  ungroup()%>%
  mutate(Std.dev = sd(Obesity_cases),
         COV = round((Std.dev/Obesity_mean),digits =2))
head(Obesity4)
## To keep to the original format of the data, i assigned a sorting mechanism to an object called *reorder*, and then used the slice(match()) functions to rearrange them in my desired order. R naturally arranges rows based on numerics first, and strings second.
reorder<-c("Under 16", "16 to 24", "25 to 34", "35 to 44", "45 to 54", "55 to 64", "65 to 74", "75 & Over")
Obesity4%>% slice(match(reorder, Age_Groups))

## Scatter plot of obesity cases in different Age groups when the data is summarized using measures of central tendancy or 5 number summary.

Obesity4 %>%
  ggplot() + geom_point(mapping=aes(y=Obesity_mean, x=Age_Groups,color="red"),size=4) + geom_point(mapping = aes(y=Obesity_Min, x=Age_Groups,color="blue"),size=2) + geom_point(mapping = aes(y=Obesity_max, x=Age_Groups,color="green"), size=6) + coord_cartesian(ylim=c(40,4500)) + labs(y="Obesity Cases") + scale_colour_manual(name = 'Summary_Stat', 
         values =c('green'='green','red'='red','blue'='blue'), labels = c('Min','Max','Mean')) +
  ggtitle("Summarized Hospitalization cases with respect to Age Groups")

##Scatter plot with the standard error showing very poor correlation between coefficient of variation(COV) and average number of hospitalizations.

Obesity4 %>%
  ggplot() + geom_point(mapping=aes(y=Obesity_mean, x=COV,color=Age_Groups),size=4) + geom_smooth(aes(y=Obesity_mean, x=COV),method = "lm", se =TRUE, color = "black") + coord_cartesian(ylim=c(40,2500)) + labs(y="Average number of cases") +
  ggtitle("Summarized Hospitalization cases with respect to Age Groups")


##Scatter plot with the standard error showing very positive correlation between standard deviation and average number of hospitalizations.

Obesity4 %>% 
  ggplot() + geom_point(mapping=aes(y=Obesity_mean, x=Std.dev,color=Age_Groups),size=4) + geom_smooth(aes(y=Obesity_mean, x=Std.dev),method = "lm", se =TRUE, color = "black") + coord_cartesian(ylim=c(40,2500)) + labs(y="Average number of cases") +
  ggtitle("Summarized Hospitalization cases with respect to Age Groups")


```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
